Reanalysis Step 1: Data Preparation

### 2025-08-17 reanalysis-1-data_preparation.qmd
mice: first shot

```{r create-variables}
# Create analysis dataset with necessary transformations
df_analysis <- df_combined |>
  mutate(
    # Outcome variables
    outcome_va_vi = outcome_VA_logMAR >= 0.3,
    outcome_va_vi_binary = if_else(outcome_va_vi, 1, 0),
    outcome_va_blind = outcome_VA_logMAR >= 1,
    outcome_va_blind_binary = if_else(outcome_va_blind, 1, 0),

    # Exposure variable
    ever_lapse_binary = if_else(person_ever_lapse == TRUE, 1, 0),

    # Demographic factors
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),
    age_cat = cut(age, breaks = c(0, 20, 45, 65, Inf), 
                  labels = c("0-20", "21-45", "46-65", "65+")),

    # DR severity (0 = No DR, 1 = NPDR, 2 = PDR)
    person_dr = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" | NPDR == "Moderate" | NPDR == "Severe" ~ 1,
      PDR == "Present" ~ 2
    ),

    # DR severity detailed (for stratified analyses)
    person_dr_severity = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 2,
      NPDR == "Severe" ~ 3,
      PDR == "Present" ~ 4
    ),

    # DR severity factor for interaction models
    dr_severity = factor(person_dr,
                        levels = c(0, 1, 2),
                        labels = c("No_DR", "NPDR", "PDR")),

    # Treatment type categorization
    treatment_type = case_when(
      anti_VEGF == 1 ~ "anti_VEGF",
      PRP_flag == 1 ~ "PRP",
      other_inject == 1 | focal_laser_flag == 1 ~ "other_treatment",
      TRUE ~ "no_treatment"
    ),
    treatment_type = factor(treatment_type, 
                           levels = c("no_treatment", "anti_VEGF", "PRP", "other_treatment")),

    # Binary treatment indicator
    any_treatment = as.numeric(anti_VEGF == 1 | PRP_flag == 1 | 
                              other_inject == 1 | focal_laser_flag == 1),

@@

    # Treatment indicators as factors
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag)
  )
```

>>>>

```{r create-variables}
# Step 1: Create person_dr first (needed for imputation)
df_analysis <- df_combined |>
  create_person_dr()  # Creates person_dr from No_DR/NPDR/PDR

# Step 2: Create other core variables
df_analysis <- df_analysis |>
  mutate(
    # Exposure variable
    ever_lapse_binary = if_else(person_ever_lapse == TRUE, 1, 0),

    # Demographic factors (keep as factors for models)
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),

@@

    # Treatment indicators as factors
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag)
  )

# Step 3: Create derived variables using transformation functions
df_analysis <- df_analysis |>
  create_outcome_variables() |>  # Creates outcome_va_vi_binary from logMAR
  create_age_cat() |>  # Creates age_cat from continuous age
  create_treatment_type() |>  # Creates treatment_type and any_treatment
  create_dr_severity_variables()  # Creates person_dr_severity and dr_severity factor
```
